---
title: "Exercise Session 0"
author: "Justin Noel"
date: "10/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

## Introduction to R

R is a programming language designed for statistical computing. It is designed for working with data and has an extensive collection of libraries for this purpose. A huge part of the value of a particular programming language is the existing collection of libraries.

I personally find R's programming syntax to be a bit strange, but I expect that we will learn to love it together. 

### R as a calculator

R is often run in interactive mode. This means that there is an interpreter running in a loop; it reads your command, executes, and then displays the output. This is much like a calculator. 

In particular we can use R as a calculator:
```{r}
sin(3)
```

```{r}
2^32
```

### Data types

Of course there are numeric types (integers, floating point numbers). 
```{r}
3L; typeof(3L);
3; typeof(3)
```
Here I executed multiple commands by separating them with a semicolon.

There are character/string types for storing text:
```{r}
"this is text"; typeof("this is text")
```

Single and double quotes are interpreted the same:
```{r}
'this' == "this"
```
But it is convenient to use them when you need to text with quotation marks:
```{r}
'these are "quotes"'
```

#### Vectors
Vectors are used for lists of elements of the same type:
```{r}
x <- c(1,3,5); x; typeof(x); x[2]+1
```
Note that we started the value of the vector into a variable x via the assignment operation <- . One can also use the = expression for assignment, but this can only be used in a simple assignment expression and may not be compatible with old versions of S-plus. I generally consider restricting assignment to simple expressions to be a good thing. 

We also accessed the second element of our vector by x[2]. Note that most programming languages start the indexing at 0, but not R. It starts at 1. This is probably more intuitive for most people, but confusing to those who have programmed with other languages.

Vectors of mixed types are coerced into one type:
```{r}
y <- c(1,"a"); y; typeof(y)
```
We can see that R decided that I did not really want to include the number 1, in my list and turned it into a string. This is both 'helpful' and confusing. It can lead to unexpected errors:
```{r}
y[1]+1
```
We can also generate vector sequences easily:
```{r}
seq1 <- 1:5; seq1; typeof(seq1)
```
```{r}
seq2 <- seq(0,10,2); seq2
```

#### Lists
To put mixed data types together we can create a list:
```{r}
list1 <- list(1, 2, "three", c(1,2,3)); list1; typeof(list)
```
We can access the elements as with vectors:
```{r}
list1[3]; typeof(list1[3])
```
Note that indexing still returns a list.
```{r}
list1[2]+2
```
Instead we need:
```{r}
list1[[2]]+2
```
### Matrices / Arrays

We can build a matrix from a vector
```{r}
M1 = matrix( 1:20, nrow = 4, ncol = 5); M1
```
Transpose operation:
```{r}
t(M1)
```

Matrix multiplication is done as follows:
```{r}
t(M1)%*%M1
```
The standard multiplication is componentwise:
```{r}
1:5*2:6; 1:5 %*% 2:6
```
Note that R tries to be 'helpful' and change the vectors into column/row vectors to make sense of the multiplication.
```{r}
1:5*3
```

We can also create arrays, which are essentially tensors:
```{r}
T1 <- array(1:24, dim = c(2,3,4)); T1
```

We can access entries in arrays/matrices in the expected way:
```{r}
M1[[2,3]]; T1[[2,1,3]];
```

### Factors
In data analysis we often have non-numeric values. These correspond to factors in R.
```{r}
genders <- c('male', 'female', 'female', 'male'); factor(genders);
```

### Data Frames
One of the most common structures in data science is the data frame. This can be thought of as a large table, like a spreadsheet, but all of the columns must be of the same type.


```{r}
participants <- data.frame(
  gender = c("male", "male", "female", "male"),
  first_name = c("John", "Moritz", "Martha", "Bernd"),
  age = c(19, 20, 18, 30)
);
print(participants)
```
We can access columns as follows:
```{r}
participants$first_name
```
Note that R tried to be 'helpful' and coerced all of the strings into factors. This is not desirable for the names.
```{r}
participants <- data.frame(
  gender = c("male", "male", "female", "male"),
  first_name = c("John", "Moritz", "Martha", "Bernd"),
  age = c(19, 20, 18, 30), stringsAsFactors = FALSE
);
print(participants)
```
We do want the gender to be factor names:
```{r}
participants$gender = factor(participants$gender); participants
```

We can get columns and rows:
```{r}
participants[1:2]; participants[1:2,]
```

We can also select from the frame:
```{r}
subset(participants, gender == "male" & age < 25)
```

## Probability functions
R has precisely 3 gazillion stats functions. Let's play with some standard ones.

Here lets grab 20 samples from a standard normal distribution:
```{r}
y <- rnorm(20); y
```
Similarly from the uniform distribution (on the interval):
```{r}
x <- runif(20); x
```

We can then plot this:
```{r}
plot(x,y)
```

Or if want to get fancy (and we do), we can use ggplot2:
```{r}
library(ggplot2);
df <- data.frame( xvals = x, yvals = y);
ggplot(df, aes(x = xvals, y = yvals)) + geom_point();
```

Or the interactive version (just for this type of webpage):
```{r message = FALSE}
library(plotly);
df <- data.frame( xvals = x, yvals = y);
pl <- ggplot(df, aes(x = xvals, y = yvals)) + geom_point();
ggplotly(pl)
```

Okay let's get fancy. Let's plot a two variable standard Gaussian.
```{r}
x = seq(-3,3, length = 20)
y = x
f = function(x,y) { exp(-(x^2+y^2)/2)*(2*pi)^(0.5) }
g = outer(x,y,f)
contour(x,y,g)
persp(x,y,g, theta = 30, phi = 10)
```

## Playing with a data set
Let's load our library of datasets and examine one. First lets see the top of the table.
```{r}
library(ISLR)
head(Auto)
```

Now the bottom:
```{r}
tail(Auto)
```

Now a summary:
```{r}
summary(Auto)
```

The basic shape:
```{r}
dim(Auto)
```

Other parts:
```{r}
colnames(Auto)
rownames(Auto)
```
Let's examine the interactions further:
```{r}
plot(Auto$mpg, Auto$cylinders)
```

Hmm...it looks it thinks the cylinders are interpreted numerically, when certain values (7) can not happen. Let's fix this and try again:
```{r}
Auto$cylinders = as.factor(Auto$cylinders)
plot(Auto$mpg ~ Auto$cylinders, legend.text =levels(Auto$cylinders))
```

Let's see how ggplot handles this:
```{r}
p <- ggplot(Auto, aes(x = cylinders, y = mpg)) + geom_boxplot()
ggplotly(p)
```

Now a histogram:
```{r}
hist(Auto$mpg)
```

Now with ggplot
```{r}
p <- ggplot(Auto, aes(x = mpg))+geom_histogram(binwidth = 5)
ggplotly(p)
```

Check pairs:
```{r}
pairs(Auto)
```

```{r}
pairs(~ mpg + weight + horsepower, Auto)
```

Okay now the ggplot equivalent with GGally.
```{r message = FALSE}
library(GGally)
ggpairs(Auto[,-9])
p<- ggpairs(Auto[,c("mpg","weight","horsepower")])
p
ggplotly(p)
```

Looks like there is a roughly quadratic relation between mpg an d horsepower.
```{r}
plot(Auto$horsepower, Auto$mpg)
```

```{r}
lm.fit=lm(mpg ~ poly(horsepower,2), data = Auto)
summary(lm.fit)
```


