---
title: "Exercise Session 0"
author: "Justin Noel"
date: "10/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to R

R is a programming language designed for statistical computing. It is designed for working with data and has an extensive collection of libraries for this purpose. A huge part of the value of a particular programming language is the existing collection of libraries.

I personally find R's programming syntax to be a bit strange, but I expect that we will learn to love it together. 

### R as a calculator

R is often run in interactive mode. This means that there is an interpreter running in a loop; it reads your command, executes, and then displays the output. This is much like a calculator. 

In particular we can use R as a calculator:
```{r}
sin(3)
```

### Data types


```{r}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
