---
title: "Problem Session 9"
author: "Justin Noel"
date: "12/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, messsage = F)
```

## R Lab
Okay the goal of this lab is to generate ROC curves and calculate the AUC for a function that outputs probabilities. 

Load our libraries
```{r}
library(tidyverse)
```
Now let's build our functions. First let us construct the relevant dataframe
```{r}
ROC_Generator = function(probs, labels) {
  df = tibble(probabilities = probs, labels = labels) %>%
    arrange(probabilities) %>%
    mutate(cum_positive = cumsum(labels))
  num_samples = nrow(df)
  num_positive = sum(select(df, labels))
  num_negative = num_samples - num_positive
  tpr = numeric(num_samples)
  fpr = numeric(num_samples)
  for(ix in seq(1,num_samples)) {
    tpr[ix] = 1 - df[[ix,3]]/num_positive 
    fpr[ix] = 1 - (ix-df[[ix,3]])/num_negative
  }
  df = add_column(df, fpr = fpr, tpr = tpr)
  #add row for when the classifier always yields positive
  df = add_row(df, fpr = 1, tpr = 1, .before = 1)
  environment()
}
```
Load predictions and data from LogisticClassifier on MNIST dataset. We will just look at the predicted probabilities for the 0 digit.
```{r}
zero_probs = preds[,1]
zero_labels = y_test[,1]
```
Let's test this:
```{r}
log_roc = ROC_Generator(zero_probs, zero_labels)
head(log_roc$df)
```
```{r}
ggplot(data = log_roc$df, aes(x=fpr, y =tpr)) +geom_line()
```

Okay let us create the desired functions now.
```{r}
roc_object = function(probs, labels) {
  roc_generator = ROC_Generator(probs,labels)
  num_samples = length(probs)
  df = roc_generator$df

  tpr_fpr_matrix = function(num_data_points = (num_samples + 1) ) {
    mat = matrix(numeric(num_data_points*2), num_data_points, 2)
    # build sequence of points to check
    s = seq(0,1,length.out=num_data_points)
    matix = 1
    ix = 2
    # For each threshold value find the row in the data frame where the probability first is less than or equal to the threshold value
    for(threshold in s) {
      while((ix <= nrow(df)) & df[ix,1]>threshold) {
        ix = ix+1
      }
      #load fpr and tpr in from previous row (which is the last below the threshold)
      mat[matix,1]=df[[ix-1,4]]
      mat[matix,2]=df[[ix-1,5]]
      matix = matix+1
    }
    return(mat)
  }
  
  better_tpr_fpr_matrix = function() {
    return( as.matrix(select(df, fpr, tpr)))
  }
  print_graph = function(num_data_points = (num_samples + 1)) {
    mat = tpr_fpr_matrix(num_data_points)
    ggplot(aes(x=mat[,1], y = mat[,2])) + geom_line()
  }
  better_print_graph = function() {
    ggplot(data = df, aes(x=fpr, y = tpr)) + geom_line()
  }
  
  AUC = function(num_data_points = (num_samples+1)) {
    mat = better_tpr_fpr_matrix()
    area = 0
    #Trapezoid method
    for(ix in seq(2, nrow(mat))) {
      dx = abs(mat[[ix-1,1]]-mat[[ix,1]])
      area = area + 0.5 * dx * (mat[[ix-1,2]]+mat[[ix,2]])
    }
    return(area)
  }
  environment()
}
```

Let us test this.
```{r}
log_roc = roc_object(zero_probs, zero_labels)
m = log_roc$tpr_fpr_matrix()
m[5000,]
```
It looks like this doesn't work so well because so many of the probabilities are so extreme. 

Let's use a better version that is built around the probabilities we actually generate.
```{r}
n = log_roc$better_tpr_fpr_matrix()
```
That's better.
```{r}
log_roc$better_print_graph()
```

```{r}
log_roc$AUC()
```

